---
title: "World Population Density"
author: "David Jackson"
date: "3/31/2021"
output: 
  html_document: 
    fig_width: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(dlookr)
library(flextable)
```

## World Population Density

```{r}
rm(list=ls())
population_density <- read.csv("./DATA/population-density.csv")
names(population_density) <- c("Country","Code","Year","Density")
```

### Descriptive Statistics 

```{r}
  population_density %>% diagnose_numeric(Density) %>% flextable()
```

### Top Five Countries by Population Density for Year 2017

```{r}
population_density %>% filter(Year =="2017") %>% top_n(5) %>%
  ggplot() + geom_col(aes(x=reorder(Country,Density),y=Density)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Top Five Counties Ranked by Population Density(2017) ",
       x="Counties Ranked",y="Population Density",caption = "(Source: http://ourworldindata.org)")

```

###  Five Countries With the Lowest Population Density for Year 2017

```{r}
population_density %>% filter(Year =="2017") %>% top_n(-5) %>%
  ggplot() + geom_col(aes(x=reorder(Country,Density),y=Density)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Five Counties With Lowest Population Density(2017) ",
       x="Counties Ranked",y="Population Density",caption = "(Source: http://ourworldindata.org)") + coord_flip()

```

### Top Five Countries by Population Density  And Year (1961 - 2017)

```{r}
population_density %>% filter(Year =="2017") %>% top_n(5) %>% 
  select(Code) %>% left_join(population_density,by = "Code") %>% 
  ggplot() + geom_line(aes(x=Year,y=Density,col =Country)) +
  scale_y_continuous(labels = comma) + 
  labs(title = "Top Five Countries Ranked by Population Density(1961-2017",caption = "(Source: http://ourworldindata.org)")
```

###  Five Countries With the Lowest Population Density (1961 - 2017)

```{r}
population_density %>% filter(Year =="2017") %>% top_n(-5) %>% 
  select(Code) %>% left_join(population_density,by = "Code") %>% 
  ggplot() + geom_line(aes(x=Year,y=Density,col =Country)) +
  scale_y_continuous(labels = comma) + 
  labs(title = "Five Countries With Lowest Population Density(1961-2017)",caption = "(Source: http://ourworldindata.org)")
```

### Calculate Rate of Change (Percent +/-) for Each count (first/last yeAR)
"How can I calculate the percent of change in population density for each country? I'm trying to use the first and last value for each country.(Slack)"

Thaks You Matthew Bacinskas @ R4DS on Slack

```{r}
rate_of_change <-population_density %>% group_by(Country) %>%
  summarise(
    First = first(Density),
    Last = last(Density),
    Percent_Change = (Last-First)/First
  )
```

Plot Rate of Change For Top 5 and Bottom 5 Countries

```{r}
top5 <- rate_of_change %>% top_n(5)
bottom5 <- rate_of_change %>% top_n(-5)
top_bottom <- rbind(top5,bottom5)
```
```{r}
ggplot(top_bottom) + geom_col(aes(x=Country,y=Percent_Change)) +
  coord_flip() + scale_y_continuous(labels=percent) 
```

